# (PART) Algebraic Structures in R {-}

# Modular Arithmetic

We will require the package `pracma`, which you may recall from linear algebra, contains "practical math applications." It has implementations of modular arithmetic and the GCD algorithms. 
```{r,echo=TRUE,warning=FALSE}
require(pracma)
```

## Integer division and GCD

The `mod` function computes the remainder after division by $n$ `idivide` gives the quotient. Thus, for example,
```{r,echo=TRUE}
mod(68,15)
idivide(68,15)
```
Since $68 = 15 \cdot 4 + 8$.

The **greatest commond divisor GCD** of two integers is computed with the `gcd(m,n)` command. For example,
```{r,echo=TRUE}
gcd(70,42)
gcd(68,15)
```
telling us that $gcd(70,42) = 14$ and $gcd(68,15) = 1$, meaning that 68 and 15 are *relatively prime*. 

The **Extended GCD** gives a little more information that can be quite useful in this course. If $d = gcd(a,b)$ then there exist integers $s,t \in \ZZ$ such that
$$
d = s a+ t b.
$$
You can find a proof of this fact in Chapter 0. We can use it here:
```{r,echo=TRUE}
gcd(70,42,extended=TRUE)
gcd(68,15,extended=TRUE)
```
Telling us that 
$$
\begin{align*}
14 & = (-1) 70 + (2) 42 \\
1 & = (2)68 + (-9) 15
\end{align*}
$$
As we can confirm with
```{r,echo=TRUE}
2*68 - 9*15
```

## Addition Mod n: $\ZZ_n$

We can generate the finite **cyclic** group $\ZZ_n = \{0, 1, 2, \ldots, n-1\}$ easily as follows:

```{r,echo=TRUE}
n = 12
G = 0:(n-1)  # generates an integer sequence from 0 to n-1.
G
```

And then we can add all of these elements mod n and make the **Cayley table** as follows: 
```{r,echo=TRUE}
n=6
G = 0:(n-1)
T = mod(outer(G,G,"+"),n)  # the outer sum computes a table of sums i + j
rownames(T) = G
colnames(T) = G
T
```

Even better, we can write a little function to do this.

```{r,echo=TRUE}
CyclicCayley <- function(n,op="+") {
  G = 0:(n-1)
  T = mod(outer(G,G,op),n)
  rownames(T) = G
  colnames(T) = G
  return(T) 
  }
```

Then 

```{r,echo=TRUE}
CyclicCayley(5)
```
and 

```{r,echo=TRUE}
CyclicCayley(15)
```

## Multiplication Mod n: $U(n)$

The group of **units** mod $n$ is the set
$$
U(n) = \{1 \le k < n \mid \gcd(k,n) = 1\}
$$
of integers that are relatively prime to $n$. As you show in PS1.2, $U(n)$ is group under multiplication mod $n$. That is, you show that it is closed under multiplication and that every number that is relatively prime to $n$ has a multipliciative inverse mod $n$.

The functions below will give us $\ZZ_n$ and $U_n$.

```{r,echo=TRUE}
Cyclic <- function(n) { 0:(n-1) }
Units <- function(n) { (1:(n-1))[gcd(1:(n-1),n)==1] }
```

For example,
```{r,echo=TRUE}
Cyclic(10)
Units(10)
```

We can multiply these by writing a MultTable function that uses the outer product

```{r,echo=TRUE}
MultTable <- function(G,n,op="*") {
  T = mod(outer(G,G,op),n)
  rownames(T) = G
  colnames(T) = G
  return(T)
  }
```

And we try it out:
```{r,echo=TRUE}
MultTable(Cyclic(10),10)
```

```{r,echo=TRUE}
MultTable(Units(10),10)
```

```{r,echo=TRUE}
MultTable(Units(2),2,op="*")
```

```{r,echo=TRUE}
MultTable(Units(3),3,op="*")
```

```{r,echo=TRUE}
MultTable(Units(5),5,op="*")
```

```{r,echo=TRUE}
MultTable(Units(7),7,op="*")
```
```{r,echo=TRUE}
MultTable(Units(11),11,op="*")
````

Suppose that we want to know the order of 5 in U(22). We could look at the Cayley table and keep multiplying by 5 until we get to 1. 
```{r,echo=TRUE}
MultTable(Units(22),22,op="*")
````

Or we can take the powers of 5 mod 22. Here are the first 10 powers of 5. We see that |5| = 5 in U(22).

```{r,echo=TRUE}
5^(0:10)
mod(5^(0:10),22)
````