# Matrix Groups

```{r,echo=TRUE,warning=FALSE}
require(pracma)
```

## The General Linear Group $GL_n(\RR)$

The **general linear group** $GL_n(\RR)$ is the group of all *invertible* $n \times n$ matrices with entries from the real numbers $\RR$. This is the world we lived in when we worked with square invertible matrices (and the invertible matrix theorem) in Math 236 Linear Algebra. 

Remember that R has a funny matrix multiplication symbol: `%*%`. For example,
```{r}
A = cbind(c(1,2,3),c(1,1,1),c(0,1,1))
B = cbind(c(-5,1,1),c(1,1,3),c(1,2,1))
A %*% B
B %*% A
```
illustrating that matrix multiplication is not-commutative and $GL_n(\RR)$ is a **nonabelian** group.

The **identity matrix** is the identity element of the group, and it can be defined with `diag(n)`. If you have `pracma` loaded then you can invert the matrix A with `inv(A)` otherwise, in base R, you can use `solve(A)`.

```{r}
inv(A)
zapsmall(A %*% inv(A))
diag(3) %*% A
```

## Matrix Multiplication Mod m:  $GL_n(\ZZ_m)$

You can use the `mod` command from `pracma` to do matrix multiplication mod m. 

For example, notice how the product looks very different for different $\ZZ_m$.

```{r}
A = rbind(c(1,3),c(2,7))
B = rbind(c(5,4),c(-3,-2))
A %*% B
mod(A %*% B,2)
mod(A %*% B,3)
mod(A %*% B,4)
```

