# Complex n-th roots of unity

## Cyclic subgroups of $\mathbb{C}^\ast$


Let's look at the cyclic subgroup generated by the complex number given in polar and rectangular form by
$$
z = 1.02 e^{2 \pi i /17} = 0.9511 + 0.3685 i
$$
First, here is how to define it in R.
```{r,echo=TRUE}
n = 17
ii = 0+1i
z = 1.02*exp(2*pi*ii/n)
z
```
Observe that the length of $z$ is 1.02 and its angle is $2 \pi/17$, or $1/17$ of a circle.


Now, we can plot $\langle z \rangle$ by raising it to a bunch of powers and plotting it according to its real and imaginary parts. You see that the positive powers of z cycle out to infinity and the negative powers cycle in to 0.
```{r,fig.width=6,fig.height=6,echo=TRUE}
G = z^(-50:100)
plot(Re(G),Im(G),col='gray',xlim=c(-2,2),ylim=c(-2,2),type='l',xlab="Re",ylab="Im")
grid()
points(Re(G),Im(G),col='black',pch=19)
points(Re(z),Im(z),pch=19,col='blue')
```

In this case we get an infinite cyclic group
$$
\langle z \rangle = \{ \ldots, z^{-3}, z^{-2}, z^{-1}, 1, z, z^2, z^3, \ldots \} \le \mathbb{C}^\ast,
$$
and all of these powers of $z$ are distinct elements in $\mathbb{C}^\ast


## Roots of Unity

Suppose now that we pick a point on the unit circle and do the same thing. For this first example, we will use a point that is 1/14 of the way around the circle. Notice that we are choosing $r = 1$ to get a point on the unit circle.
$$
w = e^{2 \pi i/14} = 0.9009698 + 0.4338837 i
$$

```{r,echo=TRUE}
n = 14
ii = 0+1i
w = exp(2*pi*ii/n)
w
```

When we plot $\langle w \rangle$ we get

```{r,fig.width=6.0,fig.height=6.0,echo=FALSE}
n = 14
ii = 0+1i
w = exp(2*pi*ii/n)
G = w^(0:(n-1))
tt = seq(0,2*pi,len=1000)
plot(cos(tt),sin(tt),type='l',col='red',xlim=c(-1.5,1.5),ylim=c(-1.5,1.5),xlab="Re",ylab="Im")
points(Re(G),Im(G),col='black',pch=19)
grid()
```

In this case we get a cyclic subgroup of order 14
$$
\langle w \rangle = \{1, w, w^2, \dots, w^{13}\} \le \mathbb{C}^\ast,
$$
and we really see, geometrically, why the term **cyclic** subgroup makes sense, because the numbers cycle around.

We can even see subgroups in here. For example, the next picture shows the subgroup $\langle w^2 \rangle \le \langle w \rangle$.
$$
\langle w^2 \rangle = \{1, w^2, w^4, w^6, w^8, w^{10}, w^{12}\} \le \{1, w, w^2,w^3,w^4,w^5,w^6,w^7,w^8,w^9,w^{10},w^{11},w^{12} w^{13}\} = \langle w \rangle.
$$

```{r,fig.width=6.0,fig.height=6.0,echo=FALSE}
n = 14
ii = 0+1i
w = exp(2*pi*ii/n)
G = w^(0:(n-1))
H = w^(2*(0:(n-1)))
tt = seq(0,2*pi,len=1000)
plot(cos(tt),sin(tt),type='l',col='red',xlim=c(-1.5,1.5),ylim=c(-1.5,1.5),xlab="Re",ylab="Im")
points(Re(G),Im(G),col='black',pch=19)
points(Re(H),Im(H),col='blue',pch=19)
grid()
```

In this example, the complex number $w$ has the property that
$$
w^{14} = 1.
$$
and if we take any other element fro the group it also has this property
$$
\text{ if } \quad w^j \in \langle w \rangle \quad \text{then} \quad (w^j)^{14} = w^{14 j} = (w^{14})^j = 1^j = 1.
$$
So all 14 elements of $\langle w rangle$ are solutions to the equation
$$
x^{14} = 1 \qquad \hbox{or} \qquad x^{14} - 1 = 0.
$$
Since this is a 14th degree polynomial equation, it has 14 complex solutions (according to the Fundamental Theorem of Algebra), and so these are *all* of the solutions to this equation. There are two real solutions 1 and -1 and the other 12 are purely complex. These are called the **14th roots of unity**.

### Cube Roots of Unity

The cube roots of unity are solutions to $x^3 - 1 = 0$. They are generated by the complex number that is 1/3 of the way around the circle.
$$
w  = e^{2 \pi i /3} = \cos(2 \pi/3) + \sin(2 \pi/3) i =  \frac{1}{2} + \frac{\sqrt{3}}{2} i.
$$

They generate a cyclic group of order 3:
$$
\langle w \rangle = \{ w^0 = 1, \quad w =  \frac{1}{2} + \frac{\sqrt{3}}{2} i, \quad w^2 =  \frac{1}{2} - \frac{\sqrt{3}}{2} i\}
$$
which we see here:

```{r,fig.width=6.0,fig.height=6.0,echo=FALSE}
n = 3
ii = 0+1i
w = exp(2*pi*ii/n)
G = w^(0:(n-1))
tt = seq(0,2*pi,len=1000)
plot(cos(tt),sin(tt),type='l',col='red',xlim=c(-1.5,1.5),ylim=c(-1.5,1.5),xlab="Re",ylab="Im")
points(Re(G),Im(G),col='black',pch=19)
grid()
```

### 4th Roots of Unity

The 4th roots of unity are solutions to $x^4 - 1 = 0$. They are generated by $i$:
$$
w  = e^{2 \pi i /4} = \cos(\pi/2) + \sin(\pi/2) i =  i.
$$
and 
$$
\langle i \rangle = \{1, i, -1, - i\}.
$$

They generate a cyclic group of order 4:

```{r,fig.width=6.0,fig.height=6.0,echo=FALSE}
n = 4
ii = 0+1i
w = exp(2*pi*ii/n)
G = w^(0:(n-1))
tt = seq(0,2*pi,len=1000)
plot(cos(tt),sin(tt),type='l',col='red',xlim=c(-1.5,1.5),ylim=c(-1.5,1.5),xlab="Re",ylab="Im")
points(Re(G),Im(G),col='black',pch=19)
grid()
```


### 8th Roots of Unity

The 8th roots of unity are solutions to $x^8 - 1 = 0$. They are generated by $i$:
$$
w  = e^{2 \pi i /8} = \cos(\pi/4) + \sin(\pi/4) i =  \frac{\sqrt{2}}{2} + \frac{\sqrt{2}}{2} i.
$$

They generate a cyclic group of order 8:

```{r,fig.width=6.0,fig.height=6.0,echo=FALSE}
n = 8
ii = 0+1i
w = exp(2*pi*ii/n)
G = w^(0:(n-1))
tt = seq(0,2*pi,len=1000)
plot(cos(tt),sin(tt),type='l',col='red',xlim=c(-1.5,1.5),ylim=c(-1.5,1.5),xlab="Re",ylab="Im")
points(Re(G),Im(G),col='black',pch=19)
grid()
```